\Ace: An Extensible Static Type System Inside a Python

### 2014-03-26 ace-oopsla14/string-appendix.tex
String proof reading for grammar etc.

Formally, input sanitation is the problem of ensuring that an arbitrary string
is converted into a safe form before potentially unsafe use. For example, consider
SQL injection attacks. To prevent such attacks, we might ensure that any string
used as input to a query does not contain unescaped SQL common sequences.

This appendix presents a type system called $\lcs$ for ensuring that input 
sanitation algorithms are implemented correctly with respect to use site 
specifications.

Unlike frameworks provided by general purpose languages such as Haskell and
Ruby, our type system provides a \emph{static} guarantee that input is always 
properly sanitized before use. We achieve this by moving portions of the regular
expression library into the type system. Therefore, type safety relies upon 
several closure and decidability results about regular languages.

Although Ur/Web achieves a similar safety guarantee, its type system is significantly
more complicated than the system defined below. We believe this simplicity 
demonstrates the power of extensibility. Instead of introducing new abstractions
based upon subtle analyses and requiring the adoption of new programming languages, 
extension designers may capture natural programming idioms directly. TODO-nrf this seems unsubstantiated or dangerously vague.

Finally, XDuce typechecks XML. Like our work, XDuce relies upon some properties of
regular expression to establish soundness and copmleteness results. Unlike
XDuce, our work is motivated by input sanitation and therefore considers arbitrary
strings (as opposed to tree-structured XML documents). Furthermore, our static
treatment of the ubiquitous ``filter" function for regular expression is novel.

>>>>

Formally, input sanitation is the problem of ensuring that an arbitrary string
is converted into a safe form before potentially unsafe use. For example, consider
SQL injection attacks. To prevent such attacks, we might ensure that any string
used as input to a query does not contain unescaped SQL command sequences.

This appendix presents a type system, $\lcs$, for ensuring that input 
sanitation algorithms are implemented correctly with respect to use site 
specifications.

Unlike frameworks provided by general purpose languages such as Haskell and
Ruby, our type system provides a \emph{static} guarantee that input is always 
properly sanitized before use. We achieve this by defining a typing relation
which captures idiomatic sanitation algorithms. Type safety relies upon 
several closure and decidability results about regular languages.

XDuce typechecks XML. Like our work, XDuce relies upon some properties of
regular expression to establish soundness and copmleteness results. Unlike
XDuce, our work is motivated by input sanitation and therefore considers arbitrary
strings (as opposed to tree-structured XML documents). Furthermore, our static
treatment of the ubiquitous regex filter (replace) function for regular expression is novel.

Finally, some research languages achieve similar safety guarantees via a specialized type
systems. The existence of such domain-specific type systems is suggestive, and 
we believe the simplicity of $\lcs$ demonstrates the effectiveness of
extensibility. Instead of introducing entirely new languages for each domain,
language extension developers may invest incrementally in obtaining static
guarnatees which address common mistakes.
